---
title: "Boruta trial"
author: "Thomas Van Hoey"
date: "2024-02-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(Boruta)
library(nycflights13)
library(ranger)
library(vip)
library(beepr)
library(here)
```
Should start at Boruta testing (the R project)
```{r}
here()
```


# Helpful documentation

* Intro from LADAL: https://ladal.edu.au/tree.html#3_Boruta 
* Package site: https://gitlab.com/mbq/Boruta/
* Vignette: https://cran.r-project.org/web/packages/Boruta/vignettes/inahurry.pdf 
* Paper: https://www.jstatsoft.org/article/view/v036i11 



# Iris

```{r}
set.seed(17)

irisE <-
  cbind(setNames(data.frame(apply(iris[,-5], 2, sample)),
                 sprintf("Nonsense%d", 1:4)
                 ),
        iris)


boruta_on_irisE <-
  Boruta(Species ~ ., data = irisE)

boruta_on_irisE
```

```{r}
attStats(boruta_on_irisE)
```

```{r}
Boruta::plotImpHistory(boruta_on_irisE)
```

```{r}
boruta_on_irisE %>% plot()
```



# Flights

Case study comes from https://www.tidymodels.org/start/recipes/ 


```{r}

flight_data <- 
  flights %>% 
  mutate(
    # Convert the arrival delay to a factor
    arr_delay = ifelse(arr_delay >= 30, "late", "on_time"),
    arr_delay = factor(arr_delay),
    # We will use the date (not date-time) in the recipe below
    date = lubridate::as_date(time_hour)
  ) %>% 
  # Include the weather data
  inner_join(weather, by = c("origin", "time_hour")) %>% 
  # Only retain the specific columns we will use
  select(dep_time, flight, origin, dest, air_time, distance, 
         carrier, date, arr_delay, time_hour) %>% 
  # Exclude missing data
  na.omit() %>% 
  # For creating models, it is better to have qualitative columns
  # encoded as factors (instead of character strings)
  mutate_if(is.character, as.factor) %>% 
  
  mutate(weekday = factor(weekdays(date),
                          levels = c("Monday", "Tuesday", "Wednesday",
                                     "Thursday", "Friday",
                                     "Saturday", "Sunday")))
```

```{r}
str(flight_data)
```


Let's say we want to predict whether there is a delay

arr_delay ~ dep_time + origin + dest + air_time + distance + carrier + date + weekday

## Random forest (ranger)

```{r}
rf1 <-
  ranger(arr_delay ~ dep_time + origin + dest + air_time + distance + carrier + date + weekday,
         data = flight_data,
         importance = "impurity_corrected") # don't forget to put this


beep()

rf1
```


```{r}
vip(rf1)
```

```{r}
rf1_predictions <-
  predict(rf1, data = flight_data)
```

For prediction better without variable importance, so we run again without that variable

```{r}
rf2 <-
    ranger(arr_delay ~ dep_time + origin + dest + air_time + distance + carrier + date + weekday,
         data = flight_data)
```


```{r}
rf2_predictions <-
  predict(rf2, data = flight_data)

# rf2_predictions$predictions
```

```{r}
rf1_pred_df <-
  bind_cols(flight_data, .pred = rf2_predictions$predictions)

Hmisc::somers2(as.numeric(rf1_pred_df$.pred) - 1, as.numeric(rf1_pred_df$arr_delay) - 1) %>% 
  enframe() %>% 
  mutate(value = round(value, 2))
```


## Boruta

The following will take a long time to run (like it was more 35 minutes I think),
so best to load in the object I saved it in.


```{r}
#### DONT' RUN BUT LOAD THE RDS ####

# boruta2 <-
#   Boruta(arr_delay ~ dep_time + origin + dest + air_time + distance + carrier + date + weekday,
#          data = flight_data)
# 
# beep()
# 
# 
# write_rds(boruta2,
#           file = "/Users/Thomas/Desktop/boruta2_model.rds")

boruta2_model <-
  read_rds(here("boruta2_model.rds"))
```

The message the the Boruta algorithm spit out

``` 
Growing trees.. Progress: 37%. Estimated remaining time: 51 seconds.
Growing trees.. Progress: 79%. Estimated remaining time: 16 seconds.
Computing permutation importance.. Progress: 31%. Estimated remaining time: 1 minute, 8 seconds.
Computing permutation importance.. Progress: 62%. Estimated remaining time: 38 seconds.
Computing permutation importance.. Progress: 95%. Estimated remaining time: 4 seconds.
Growing trees.. Progress: 38%. Estimated remaining time: 50 seconds.
Growing trees.. Progress: 78%. Estimated remaining time: 17 seconds.
Computing permutation importance.. Progress: 29%. Estimated remaining time: 1 minute, 14 seconds.
Computing permutation importance.. Progress: 56%. Estimated remaining time: 47 seconds.
Computing permutation importance.. Progress: 87%. Estimated remaining time: 13 seconds.
Growing trees.. Progress: 32%. Estimated remaining time: 1 minute, 5 seconds.
Growing trees.. Progress: 65%. Estimated remaining time: 34 seconds.
Growing trees.. Progress: 97%. Estimated remaining time: 3 seconds.
Computing permutation importance.. Progress: 24%. Estimated remaining time: 1 minute, 36 seconds.
Computing permutation importance.. Progress: 50%. Estimated remaining time: 1 minute, 2 seconds.
Computing permutation importance.. Progress: 74%. Estimated remaining time: 33 seconds.
Computing permutation importance.. Progress: 95%. Estimated remaining time: 6 seconds.
Growing trees.. Progress: 31%. Estimated remaining time: 1 minute, 7 seconds.
Growing trees.. Progress: 66%. Estimated remaining time: 32 seconds.
Growing trees.. Progress: 98%. Estimated remaining time: 2 seconds.
Computing permutation importance.. Progress: 28%. Estimated remaining time: 1 minute, 21 seconds.
Computing permutation importance.. Progress: 50%. Estimated remaining time: 1 minute, 1 seconds.
Computing permutation importance.. Progress: 74%. Estimated remaining time: 31 seconds.
Computing permutation importance.. Progress: 96%. Estimated remaining time: 5 seconds.
Growing trees.. Progress: 36%. Estimated remaining time: 55 seconds.
Growing trees.. Progress: 69%. Estimated remaining time: 27 seconds.
Computing permutation importance.. Progress: 28%. Estimated remaining time: 1 minute, 21 seconds.
Computing permutation importance.. Progress: 54%. Estimated remaining time: 51 seconds.
Computing permutation importance.. Progress: 82%. Estimated remaining time: 20 seconds.
Growing trees.. Progress: 36%. Estimated remaining time: 56 seconds.
Growing trees.. Progress: 69%. Estimated remaining time: 27 seconds.
Computing permutation importance.. Progress: 24%. Estimated remaining time: 1 minute, 37 seconds.
Computing permutation importance.. Progress: 49%. Estimated remaining time: 1 minute, 5 seconds.
Computing permutation importance.. Progress: 74%. Estimated remaining time: 31 seconds.
Growing trees.. Progress: 36%. Estimated remaining time: 56 seconds.
Growing trees.. Progress: 73%. Estimated remaining time: 23 seconds.
Computing permutation importance.. Progress: 30%. Estimated remaining time: 1 minute, 11 seconds.
Computing permutation importance.. Progress: 58%. Estimated remaining time: 44 seconds.
Computing permutation importance.. Progress: 89%. Estimated remaining time: 11 seconds.
Growing trees.. Progress: 34%. Estimated remaining time: 1 minute, 0 seconds.
Growing trees.. Progress: 73%. Estimated remaining time: 22 seconds.
Computing permutation importance.. Progress: 30%. Estimated remaining time: 1 minute, 13 seconds.
Computing permutation importance.. Progress: 60%. Estimated remaining time: 40 seconds.
Computing permutation importance.. Progress: 89%. Estimated remaining time: 11 seconds.
Growing trees.. Progress: 39%. Estimated remaining time: 48 seconds.
Growing trees.. Progress: 79%. Estimated remaining time: 16 seconds.
Computing permutation importance.. Progress: 27%. Estimated remaining time: 1 minute, 24 seconds.
Computing permutation importance.. Progress: 56%. Estimated remaining time: 48 seconds.
Computing permutation importance.. Progress: 81%. Estimated remaining time: 21 seconds.
Growing trees.. Progress: 35%. Estimated remaining time: 57 seconds.
Growing trees.. Progress: 71%. Estimated remaining time: 25 seconds.
Computing permutation importance.. Progress: 25%. Estimated remaining time: 1 minute, 31 seconds.
Computing permutation importance.. Progress: 47%. Estimated remaining time: 1 minute, 9 seconds.
Computing permutation importance.. Progress: 75%. Estimated remaining time: 31 seconds.
Computing permutation importance.. Progress: 98%. Estimated remaining time: 2 seconds.
```


```{r}
print(boruta2_model)
```

```{r}
boruta2_model$finalDecision
```

```{r}
# extract decision
getConfirmedFormula(boruta2_model)
```


```{r}
plotImpHistory(boruta2_model)
```

Based on this, we don't throw it out.

```{r}
boruta2_model %>% plot()
```

```{r}
borutacategory <-
  attStats(boruta2_model) %>% 
  rownames_to_column(var = "variable") %>% 
  select(variable, decision)

borutacategory
```


```{r}
boruta2_model$ImpHistory %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "iteration") %>% 
  select(-starts_with("shadow")) %>% 
  pivot_longer(-iteration,
               names_to = "variable",
               values_to = "value") %>% 
  left_join(borutacategory %>% select(variable, decision),
            by = join_by(variable)) %>% 

  ggplot() +
  geom_boxplot(aes(x = reorder(variable, - value), value, fill = decision)) # color can be changed
```



```{r}
boruta2_model %>% summary()
```


