---
title: "Genitives and Datives"
author: "Thomas Van Hoey"
date: "2024-02-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(Boruta)
library(Hmisc)
library(janitor)
library(beepr)

set.seed(1234)
```


This is still a test run for the Boruta package.

We're going to remake a study about genitives and datives.

Szmrecsanyi, Benedikt, Jason Grafmiller, Joan Bresnan, Anette Rosenbach, Sali Tagliamonte & Simon Todd. 2017. Spoken syntax in a comparative perspective: The dative and genitive alternation in varieties of English. Glossa: a journal of general linguistics 2(1). 86. https://doi.org/10.5334/gjgl.310.

With the Boruta package, we can have a better grip on variable importance.

We follow the same predictors etc. as in that paper:

> We conducted an independent CRF run for each of the varieties, per alternation

Note:

> permutation variable importance [rather than impurity or impurity_corrected]


# Genitive

```{r}
gen_orig <-
  read_csv(here("genitives-finalized.csv")) %>% 
  clean_names() %>% 
  mutate(across(c(response_variable, possessor_animacy, possessor_definiteness,
                  possessor_length, possessum_length, semantic_relation,final_sibilancy),
                ~ factor(.x)))
```

```{r}
skimr::skim(gen_orig)
```

Because `persistence` is missing from a number of cases, we don't include it. 
It's also not reported in the paper, even though it's one of the predictors.


Four subsets
```{r}
gen_can <- filter(gen_orig, variety == "CAN")
gen_nz <- filter(gen_orig, variety == "NZ")
gen_uk <- filter(gen_orig, variety == "UK")
gen_us <- filter(gen_orig, variety == "US")
```


Genitive formula
``` 
response_variable ~ possessor_animacy + possessor_definiteness + possessor_length + possessum_length + semantic_relation + final_sibilancy
```

## Gen Boruta Canada

```{r}
boruta_gen_can <-
  Boruta(response_variable ~ possessor_animacy + possessor_definiteness + possessor_length +
           possessum_length + semantic_relation + final_sibilancy,
         data = gen_can)
```

```{r}
boruta_gen_can

attStats(boruta_gen_can)

plotImpHistory(boruta_gen_can)

boruta_gen_can %>% plot()

boruta_gen_can$finalDecision %>% as.data.frame()
```

```{r}
boruta_gen_can_decision <-
  attStats(boruta_gen_can) %>% 
  rownames_to_column(var = "variable") %>% 
  select(variable, decision)
```


```{r}
boruta_gen_can$ImpHistory %>% 
  as.data.frame() %>%
  rownames_to_column(var = "iteration") %>% 
  select(-starts_with("shadow")) %>% 
  pivot_longer(-iteration,
               names_to = "variable",
               values_to = "value") %>% 
  left_join(boruta_gen_can_decision,
            by = join_by(variable)) %>% 
  ggplot() +
  geom_boxplot(aes(x = reorder(variable, value), 
                   value,
                   fill = decision)) + # color can be changed
  coord_flip()
```

## Gen Boruta US

```{r}
boruta_gen_us <-
  Boruta(response_variable ~ possessor_animacy + possessor_definiteness + possessor_length +
           possessum_length + semantic_relation + final_sibilancy,
         data = gen_us)
```

```{r}
boruta_gen_us

attStats(boruta_gen_us)

plotImpHistory(boruta_gen_us)

boruta_gen_us %>% plot()

boruta_gen_us$finalDecision %>% as.data.frame()
```

```{r}
boruta_gen_us_decision <-
  attStats(boruta_gen_us) %>% 
  rownames_to_column(var = "variable") %>% 
  select(variable, decision)
```

```{r}
boruta_gen_us$ImpHistory %>% 
  as.data.frame() %>%
  rownames_to_column(var = "iteration") %>% 
  select(-starts_with("shadow")) %>% 
  pivot_longer(-iteration,
               names_to = "variable",
               values_to = "value") %>% 
  left_join(boruta_gen_us_decision,
            by = join_by(variable)) %>% 
  ggplot() +
  geom_boxplot(aes(x = reorder(variable, value), 
                   value,
                   fill = decision)) + # color can be changed
  coord_flip()
```

## Gen Boruta NZ

```{r}
boruta_gen_nz <-
  Boruta(response_variable ~ possessor_animacy + possessor_definiteness + possessor_length +
           possessum_length + semantic_relation + final_sibilancy,
         data = gen_nz)
```

```{r}
boruta_gen_nz

attStats(boruta_gen_nz)

plotImpHistory(boruta_gen_nz)

boruta_gen_nz %>% plot()

boruta_gen_nz$finalDecision %>% as.data.frame()
```

```{r}
boruta_gen_nz_decision <-
  attStats(boruta_gen_nz) %>% 
  rownames_to_column(var = "variable") %>% 
  select(variable, decision)
```

```{r}
boruta_gen_nz$ImpHistory %>% 
  as.data.frame() %>%
  rownames_to_column(var = "iteration") %>% 
  select(-starts_with("shadow")) %>% 
  pivot_longer(-iteration,
               names_to = "variable",
               values_to = "value") %>% 
  left_join(boruta_gen_nz_decision,
            by = join_by(variable)) %>% 
  ggplot() +
  geom_boxplot(aes(x = reorder(variable, value), 
                   value,
                   fill = decision)) + # color can be changed
  coord_flip()
```

**Possessum_length deemed unimportant!!!**


## Gen Boruta UK

```{r}
boruta_gen_uk <-
  Boruta(response_variable ~ possessor_animacy + possessor_definiteness + possessor_length +
           possessum_length + semantic_relation + final_sibilancy,
         data = gen_uk)
```

```{r}
boruta_gen_uk

attStats(boruta_gen_uk)

plotImpHistory(boruta_gen_uk)

boruta_gen_uk %>% plot()

boruta_gen_uk$finalDecision %>% as.data.frame()
```

```{r}
boruta_gen_uk_decision <-
  attStats(boruta_gen_uk) %>% 
  rownames_to_column(var = "variable") %>% 
  select(variable, decision)
```

```{r}
boruta_gen_uk$ImpHistory %>% 
  as.data.frame() %>%
  rownames_to_column(var = "iteration") %>% 
  select(-starts_with("shadow")) %>% 
  pivot_longer(-iteration,
               names_to = "variable",
               values_to = "value") %>% 
  left_join(boruta_gen_uk_decision,
            by = join_by(variable)) %>% 
  ggplot() +
  geom_boxplot(aes(x = reorder(variable, value), 
                   value,
                   fill = decision)) + # color can be changed
  coord_flip()
```

# Dative


```{r}
dat_orig <-
  read_csv(here("datives-finalized.csv")) %>% 
  clean_names() %>% 
  mutate(across(c(response_variable, 
                  recipient_type, recipient_animacy, recipient_definiteness,
                  theme_type, theme_animacy, theme_definiteness,
                  semantics),
                ~ factor(.x))) %>% 
  mutate(l_diff = log(recipient_length) - log(theme_length))
```

```{r}
skimr::skim(dat_orig)
```


Four subsets
```{r}
dat_can <- filter(dat_orig, variety == "CAN")
dat_nz <- filter(dat_orig, variety == "NZ")
dat_uk <- filter(dat_orig, variety == "UK")
dat_us <- filter(dat_orig, variety == "US")
```


datitive formula
``` 
response_variable ~ 
  recipient_type + recipient_animacy + recipient_definiteness +
  theme_type + theme_animacy +  theme_definiteness +
  l_diff + semantics
```

## dat Boruta Canada

```{r}
boruta_dat_can <-
  Boruta(response_variable ~ recipient_type + recipient_animacy +
           recipient_definiteness + theme_type + theme_animacy +
           theme_definiteness + l_diff + semantics,
         data = dat_can)
```

```{r}
boruta_dat_can

attStats(boruta_dat_can)

plotImpHistory(boruta_dat_can)

boruta_dat_can %>% plot()

boruta_dat_can$finalDecision %>% as.data.frame()
```

```{r}
boruta_dat_can_decision <-
  attStats(boruta_dat_can) %>% 
  rownames_to_column(var = "variable") %>% 
  select(variable, decision)
```


```{r}
boruta_dat_can$ImpHistory %>% 
  as.data.frame() %>%
  rownames_to_column(var = "iteration") %>% 
  select(-starts_with("shadow")) %>% 
  pivot_longer(-iteration,
               names_to = "variable",
               values_to = "value") %>% 
  left_join(boruta_dat_can_decision,
            by = join_by(variable)) %>% 
  ggplot() +
  geom_boxplot(aes(x = reorder(variable, value), 
                   value,
                   fill = decision)) + # color can be changed
  coord_flip()
```

## dat Boruta US

```{r}
boruta_dat_us <-
  Boruta(response_variable ~ recipient_type + recipient_animacy +
           recipient_definiteness + theme_type + theme_animacy +
           theme_definiteness + l_diff + semantics,
         data = dat_us)
```

```{r}
boruta_dat_us

attStats(boruta_dat_us)

plotImpHistory(boruta_dat_us)

boruta_dat_us %>% plot()

boruta_dat_us$finalDecision %>% as.data.frame()
```

```{r}
boruta_dat_us_decision <-
  attStats(boruta_dat_us) %>% 
  rownames_to_column(var = "variable") %>% 
  select(variable, decision)
```

```{r}
boruta_dat_us$ImpHistory %>% 
  as.data.frame() %>%
  rownames_to_column(var = "iteration") %>% 
  select(-starts_with("shadow")) %>% 
  pivot_longer(-iteration,
               names_to = "variable",
               values_to = "value") %>% 
  left_join(boruta_dat_us_decision,
            by = join_by(variable)) %>% 
  ggplot() +
  geom_boxplot(aes(x = reorder(variable, value), 
                   value,
                   fill = decision)) + # color can be changed
  coord_flip()
```

## dat Boruta NZ

```{r}
boruta_dat_nz <-
  Boruta(response_variable ~ recipient_type + recipient_animacy +
           recipient_definiteness + theme_type + theme_animacy +
           theme_definiteness + l_diff + semantics,
         data = dat_nz)
```

```{r}
boruta_dat_nz

attStats(boruta_dat_nz)

plotImpHistory(boruta_dat_nz)

boruta_dat_nz %>% plot()

boruta_dat_nz$finalDecision %>% as.data.frame()
```


```{r}
boruta_dat_nz_decision <-
  attStats(boruta_dat_nz) %>% 
  rownames_to_column(var = "variable") %>% 
  select(variable, decision)
```


```{r}
boruta_dat_nz$ImpHistory %>% 
  as.data.frame() %>%
  rownames_to_column(var = "iteration") %>% 
  select(-starts_with("shadow")) %>% 
  pivot_longer(-iteration,
               names_to = "variable",
               values_to = "value") %>% 
  left_join(boruta_dat_nz_decision,
            by = join_by(variable)) %>% 
  ggplot() +
  geom_boxplot(aes(x = reorder(variable, value), 
                   value,
                   fill = decision)) + # color can be changed
  coord_flip()
```

**Theme animacy deemed unimportant**


## dat Boruta UK

```{r}
boruta_dat_uk <-
  Boruta(response_variable ~ recipient_type + recipient_animacy +
           recipient_definiteness + theme_type + theme_animacy +
           theme_definiteness + l_diff + semantics,
         data = dat_uk)
```

```{r}
boruta_dat_uk

attStats(boruta_dat_uk)

plotImpHistory(boruta_dat_uk)

boruta_dat_uk %>% plot()

boruta_dat_uk$finalDecision %>% as.data.frame()
```


```{r}
boruta_dat_uk_decision <-
  attStats(boruta_dat_uk) %>% 
  rownames_to_column(var = "variable") %>% 
  select(variable, decision)
```

```{r}
boruta_dat_uk$ImpHistory %>% 
  as.data.frame() %>%
  rownames_to_column(var = "iteration") %>% 
  select(-starts_with("shadow")) %>% 
  pivot_longer(-iteration,
               names_to = "variable",
               values_to = "value") %>% 
  left_join(boruta_dat_uk_decision,
            by = join_by(variable)) %>% 
  ggplot() +
  geom_boxplot(aes(x = reorder(variable, value), 
                   value,
                   fill = decision)) + # color can be changed
  coord_flip()
```


**Theme animacy and Recipient animacy rejected as useful**





```{r}
beep()
```











































